---
title: "forward_inverse_modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{forward_inverse_modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 4,
  fig.height = 4,
  out.width = "100%"
)
```

```{r setup}
devtools::load_all(".")
library(dplyr)
library(plotly)
library(BayesianTools)
```

This vignette walks through forward and inverse radiative transfer modelling with SABER. All examples use an optically shallow scenario with three benthic classes. Each section has a standard (`am03`) and a SICF-enabled (`am03_sicf`) variant side by side.

## Setup

Benthic classes must be loaded before the spectral cache is built. The next call to `iop_from_oac()` triggers cache initialisation at the analysis wavelength grid, after which `compute_r_rs_b_lmm()` returns $R_\mathrm{B}(\lambda)$ interpolated to those wavelengths.

```{r}
wavelength <- seq(401, 749, 5)

list_benthic_classes()
select_benthic_classes(c("Sand_2019", "Eelgrass_2019", "Mud_2019"))

oac_par <- c(
  chl        = 5.0,
  a_dg_440   = 0.9,
  bb_p_550   = 0.005,
  a_dg_s     = 0.014,
  bb_p_gamma = 0.5
)

iop <- iop_from_oac(wavelength, par = oac_par)

r_b <- compute_r_rs_b_lmm(c(
  r_rs_b_Sand_2019     = 0.55,
  r_rs_b_Eelgrass_2019 = 0.25,
  r_rs_b_Mud_2019      = 0.20
))

plot_ly(x = wavelength) %>%
  add_lines(y = iop$a,  name = "a [m⁻¹]") %>%
  add_lines(y = iop$bb, name = "bb [m⁻¹]")

plot_ly(x = wavelength) %>%
  add_lines(y = r_b, name = "R_B")
```

## Forward modelling

### Standard elastic model (`am03`)

`forward_am03()` computes sub-surface $R_\text{rs}$ from elastic scattering only. Non-NULL `h_w` and `r_b` activate the shallow-water treatment.

```{r}
rrs_elastic <- forward_am03(
  wavelength = wavelength,
  iop        = iop,
  water_type = 2,
  theta_sun  = 20,
  theta_view = 0,
  h_w        = 5,
  r_b        = r_b
)

# Add realistic noise function
add_noise <- function(rrs, snr = 200) {
  noise <- rnorm(length(rrs), mean = 0, sd = mean(rrs) / snr)
  rrs + noise
}

rrs_df <- tibble(wavelength = wavelength, rrs_0m = add_noise(rrs_elastic))

plot_ly(x = wavelength) %>%
  add_lines(y = rrs_elastic, name = "am03")
```

### SICF model (`am03_sicf`)

`forward_am03_sicf()` adds the Sun-Induced Chlorophyll Fluorescence (SICF) component on top of the elastic signal. It requires `chl`, `a_dg_443`, and geographic/temporal metadata (`lat`, `lon`, `date_time`) to compute the subsurface irradiance field that drives fluorescence. Setting `return_components = TRUE` gives access to the individual elastic and SICF terms.

```{r}
rrs_sicf_out <- forward_am03_sicf(
  wavelength        = wavelength,
  iop               = iop,
  water_type        = 2,
  theta_sun         = 20,
  theta_view        = 0,
  h_w               = 5,
  r_b               = r_b,
  chl               = oac_par["chl"],
  a_dg_443          = oac_par["a_dg_440"],
  phi_f             = 0.02,
  include_sicf      = TRUE,
  sicf_model        = "semi_analytical",
  depth_integration = TRUE,
  lat               = 48.5,
  lon               = -68.2,
  date_time         = as.POSIXct("2022-07-15 12:00:00", tz = "UTC"),
  return_components = TRUE
)

rrs_df_sicf <- tibble(wavelength = wavelength, rrs_0m = add_noise(rrs_sicf_out$rrs_total, snr = 200))

plot_ly(x = wavelength) %>%
  add_lines(y = rrs_sicf_out$rrs_elastic, name = "elastic") %>%
  add_lines(y = rrs_sicf_out$rrs_sicf,    name = "SICF") %>%
  add_lines(y = rrs_sicf_out$rrs_total,   name = "total")
```

The total spectrum from `am03_sicf` is used as the observed input for the SICF inversion below.

```{r}
rrs_df_sicf <- tibble(wavelength = wavelength, rrs_0m = rrs_sicf_out$rrs_total)
```

## Inversion

`make_inversion_params()` builds all inputs (bounds, prior, fixed parameters, forward model selection) from a compact set of arguments. The `sicf` flag is the only switch needed to move between the two model families — everything else is resolved automatically.

### Standard inversion (`am03`)

```{r}
cfg <- make_inversion_params(
  depth_m = 5,
  rrs_df        = rrs_df,
  par_inv       = c("chl", "a_dg_440", "bb_p_550", "a_dg_s", "bb_p_gamma"
                   ),
  water_type    = 2,
  theta_sun     = 20,
  sicf          = FALSE, force_shallow = T
)
```

**Gradient-based MAP**

```{r}
result_grad <- inverse_gradient(
  rrs           = rrs_df,
  forward_model = cfg$fwd_model,
  objective_fct = "log-ll",
  optim_mtd     = #"levenberg-marqardt",
                  "L-BFGS-B",
  par_inversed  = cfg$par_inv,
  par_fixed     = cfg$par_fixed,
  lower_b       = cfg$lower_b,
  upper_b       = cfg$upper_b,
  init_val      = cfg$init_val,
  log_prior_fn  = cfg$log_prior_fn,
  verbose  = FALSE
)
print(format(result_grad, scientific = F))
```

**MCMC**

```{r}
result_mcmc <- inverse_mcmc(
  rrs           = rrs_df,
  forward_model = cfg$fwd_model,
  par_inversed  = cfg$par_inv,
  prior         = cfg$bt_prior,
  par_fixed     = cfg$par_fixed,
  iterations    = 30000,
  burnin        = 10000,
  sampler       = "DEzs"
)
print(format(result_mcmc, scientific = F))
```

**L2 products**

```{r}
l2 <- compute_l2_products(
  rrs_df      = rrs_df,
  cfg         = cfg,
  result_grad = result_grad,
  result_mcmc = result_mcmc,
  compute_iop = TRUE,
  compute_kd  = TRUE,
  plot_rrs    = TRUE,
  plot_iop    = TRUE
)

l2$gof$grad
l2$gof$mcmc
l2$plots$rrs_recon
l2$plots$iop
```

### SICF inversion (`am03_sicf`)

The only differences from the standard case: `sicf = TRUE` and the geographic/temporal metadata required to drive the irradiance model. `phi_f` (fluorescence quantum yield) is automatically added to `par_inv` by `make_inversion_params()`.

```{r eval=FALSE}
cfg_sicf <- make_inversion_params(
  depth_m = 5,
  rrs_df            = rrs_df_sicf,
  par_inv           = c("chl", "a_dg_440", "bb_p_550", "a_dg_s", "bb_p_gamma"
                   ),
  water_type        = 2,
  theta_sun         = 20,
  sicf              = TRUE,
  sicf_model        = "semi_analytical",
  depth_integration = FALSE,
  lat               = 48.5,
  lon               = -68.2,
  date_time         = as.POSIXct("2022-07-15 12:00:00", tz = "UTC"),
  force_shallow     = TRUE
)

result_grad_sicf <- inverse_gradient(
  rrs           = rrs_df_sicf,
  forward_model = cfg_sicf$fwd_model,
  objective_fct = "log-ll",
  optim_mtd     = "levenberg-marqardt",
  par_inversed  = cfg_sicf$par_inv,
  par_fixed     = cfg_sicf$par_fixed,
  lower_b       = cfg_sicf$lower_b,
  upper_b       = cfg_sicf$upper_b,
  init_val      = cfg_sicf$init_val,
  log_prior_fn  = cfg_sicf$log_prior_fn
)

result_mcmc_sicf <- inverse_mcmc(
  rrs           = rrs_df_sicf,
  forward_model = cfg_sicf$fwd_model,
  par_inversed  = cfg_sicf$par_inv,
  prior         = cfg_sicf$bt_prior,
  par_fixed     = cfg_sicf$par_fixed,
  iterations    = 35000,
  burnin        = 12000,
  sampler       = "DEzs"
)

l2_sicf <- compute_l2_products(
  rrs_df      = rrs_df_sicf,
  cfg         = cfg_sicf,
  result_grad = result_grad_sicf,
  result_mcmc = result_mcmc_sicf,
  compute_iop = TRUE,
  compute_kd  = TRUE,
  plot_rrs    = TRUE,
  plot_iop    = TRUE
)

l2_sicf$gof$grad
l2_sicf$gof$mcmc
l2_sicf$plots$rrs_recon
l2_sicf$plots$iop
```

The returned `l2` / `l2_sicf` lists have the same structure in both cases: `rrs_recon`, `gof` (with `rmse`, `mape`, `r2`), `iop`, `Kd`, and ggplot2 `plots` — all split into `$grad` and `$mcmc` sub-lists.

