---
title: "forward_inverse_modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{forward_inverse_modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
load_all()
library(SABER)
library(readr)
library(dplyr)
library(plotly)
library(BayesianTools)
```

This vignette teaches the basics of forward and inverse radiative tranfer modelling with SABER.

## Forward modelling

We'll use the Albert & Mobley 2003 (`"am03"`) forward model to generate a $R_\text{rs}$ spectrum from Optically Active Constituent (OAC). To compute the $R_\text{rs}$, a forward model needs the IOP. If we have spectral IOP we could pass them directly to the forward model. Instead, here, we will compute the IOP from some OAC values. 

### Computing the IOP from the OAC

The first step is to choose a bio-optical model to compute the Inherent Optical Proprieties (IOP) from the OAC. To do that, SABER offer the `iop_from_oac` function. In this case we pass the spectral slope with the OAC. *Details different options*
The `verbose = T` argument will create plots of the retrieved spectral IOP, this way we can quickly check what's going on. 

```{r}
wavelength <- seq(400, 700, 5)
# Without manual slope
# oac <- tibble(
#   name = c("chl", "a_g_440", "bb_p_550"),
#   value = c(2.8, 4.3, 0.004)
#   )
# With manual slope
oac <- tibble(
  name = c("chl", "a_g_440", "bb_p_550", "a_g_s_g", "a_g_s_d", "bb_p_gamma"),
  value = c(5, 0.9, 0.005, 0.014, 0.003, 0.5)
)

oac_vec <- stats::setNames(oac$value, oac$name)

iop <- iop_from_oac(
  wavelength,
  par = oac_vec
  #rrs = NULL,
  #optically_shallow = T,
  #verbose = T
)
```

### Runing the forward model

To run the forward model we will create a list that contains the inputs the model need. 
The model we use here is Albert & Mobley 2003 (`"am03"`).

```{r}
test_case <- tibble(
  name = c("water_type", "theta_view", "theta_sun"),
  value = c(2, 0, 20)
)

test_case_vec <- stats::setNames(test_case$value, test_case$name)

rrs <- forward_am03(
  wavelength = wavelength,
  iop = iop,
  water_type = test_case_vec["water_type"],
  theta_view = test_case_vec["theta_view"],
  theta_sun = test_case_vec["theta_sun"],

  h_w = NULL,
  r_b = NULL
)
rrs_vec <- tibble(
  wavelength = wavelength,
  rrs_0m = rrs
)

ply <- plot_ly(rrs_vec) %>%
  add_lines(
    x = ~wavelength,
    y = ~rrs_0m
  )

ply
```

## Inversion

The inversion process consist in finding the optimal input parameter to a forward model, that will provide the best solution to an objective function. An objective function compares the observed $R_\text{rs}$ with the modeled $\hat{R_\text{rs}}$. Objective function includes residual sum of square (rss), log-likelihood (log-ll) or the spectral error index of Lee 1999 (lee99).

Two categories of inversion are implemented in SABER, gradient based, and markov-chain monte-carlo (mcmc).
In the following section we will demonstrate the use of gradient based inversion with augmented lagrangian optimization.

### MCMC inversion

Using uniform prior for ease of illustration (pass NULL in bayesconfig prior). Only retrieved opticall deep components.

```{r}
par_fixed <- c(
  "water_type" = 2,
  "theta_view" = 0,
  "theta_sun" = 20,
  "h_w" = NULL,
  "r_rs_b_saccharina" = NULL,
  "r_rs_b_cca" = NULL,
  "r_rs_b_gravel" = NULL
  #, sd = 0.5
)


par_to_inverse = c(#"theta_sun",
  "chl", "a_g_440","a_g_s",
  #"a_g_s_d",
  "bb_p_550", "bb_p_gamma",

 # "a_nap_440",  "a_nap_sd",

  # "h_w",  "r_rs_b_saccharina" , "r_rs_b_cca" , "r_rs_b_gravel" ,

  "sd"
)

lower = c(0.5, 0.1, 0.0008, #0.0002,
          0.002, 0.2,
          # 0.5, 0.1, 0.1, 0.1,
          0.0001
)

best = c(3, 0.75, 0.017, 0.005, 0.46,
         #  4, 0.5, 0.1, 0.25,
         0.1
)

upper = c(30, 1.5, 0.020, #0.002,
          0.015, 1,
          #  12, 1, 1, 1,
          10
)

test_inverse = inverse_mcmc(
  rrs = rrs_vec,
  forward_model = "am03",
  par_inversed = par_to_inverse,
  prior = NULL,
  lower = lower,
  best = NULL,
  upper = upper,
  par_fixed = par_fixed,
  iterations = 30000,
  burnin = 5000,
  sampler = "DEzs"
)
print(test_inverse)

```

```{r}
oac_mcmc <-test_inverse[grep("sd$", names(test_inverse), invert = T)]

iop_mcmc <- iop_from_oac(
  rrs_vec$wavelength,
  oac_mcmc
)

rrs_mcmc <- forward_am03(
  wavelength = wavelength,
  iop = iop_mcmc,
  water_type = test_case_vec["water_type"],
  theta_view = test_case_vec["theta_view"],
  theta_sun = test_case_vec["theta_sun"],
  
  h_w = NULL,
  r_b = NULL
)

ply <- plot_ly() %>%
  add_lines(x = rrs_vec$wavelength, y = rrs_vec$rrs_0m, name = "rrs_input") %>%
  add_lines(x = rrs_vec$wavelength, y = rrs_mcmc, name = "rrs_retrieved")

ply
```

