---
title: "forward_inverse_modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{forward_inverse_modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#load_all()
library(SABER)
library(readr)
library(dplyr)
library(plotly)
library(BayesianTools)
```

This vignette teaches the basics of forward and inverse radiative transfer modelling with SABER.

## Forward modelling

The SA paramterized forward model by Albert & Mobley 2003 (`"am03"`) is used here as the base RT forward model (tweaked for spectral slopes and bio-optical formulation) to generate spectral $R_\text{rs}$ from user-driven Optically Active Constituent (OAC). The forward modeling is seperated into two steps, first the OAC are converted to Inherent Optical Properties (IOP), then the IOP (and benthic information: $H$ and $R_\mathrm{B}$, where applicable) are used as input to the RT model to generate $R_\text{rs}$.

### Computing the IOP from the OAC

The first step is to choose a bio-optical model to compute the Inherent Optical Proprieties (IOP) from the OAC. To do that, SABER offer the `iop_from_oac` function. In this case we pass the spectral slopes with the OAC in a named vector `par`. *Details different options*

```{r}
wavelength <- seq(400, 700, 5)
# Without manual slope
# oac <- tibble(
#   name = c("chl", "a_g_440", "bb_p_550"),
#   value = c(2.8, 4.3, 0.004)
#   )
# With manual slope
oac <- tibble(
  name = c("chl", "a_g_440", "bb_p_550", "a_g_s_g", "a_g_s_d", "bb_p_gamma"),
  value = c(5, 0.9, 0.005, 0.014, 0.003, 0.5)
)

oac_vec <- stats::setNames(oac$value, oac$name)

iop <- iop_from_oac(
  wavelength,
  par = oac_vec
)

iop_vec = tibble(wavelength = wavelength) %>%
  bind_cols(as_tibble(iop))

ply <- plot_ly(iop_vec) %>%
  add_lines(
    x = ~wavelength,
    y = ~a,
    name = "at_w"
  ) %>%
  add_lines(
    x = ~wavelength,
    y = ~bb,
    name = "bb"
  ) 
ply
```

### Runing the forward model

Next, the obtained spectral IOPs are input into the SA RT model (`"am03"`) to generate the $R_\text{rs}$. 

```{r}
test_case <- tibble(
  name = c("water_type", "theta_view", "theta_sun"),
  value = c(2, 0, 20)
)

test_case_vec <- stats::setNames(test_case$value, test_case$name)

rrs <- forward_am03(
  wavelength = wavelength,
  iop = iop,
  water_type = test_case_vec["water_type"],
  theta_view = test_case_vec["theta_view"],
  theta_sun = test_case_vec["theta_sun"],

  h_w = NULL, #optically deep case
  r_b = NULL
)
rrs_vec <- tibble(
  wavelength = wavelength,
  rrs_0m = rrs
)

ply <- plot_ly(rrs_vec) %>%
  add_lines(
    x = ~wavelength,
    y = ~rrs_0m
  )

ply
```
### Benthic reflectance

SABER has complete support for optically shallow waters, where the benthic reflectance ($R_\mathrm{B}$) plays a crucial role. The $R_\mathrm{B}$ in SABER is calculated as a aerial fraction constrained sum of three pure benthic class spectra. The $R_\mathrm{B}$ can be calculated using `compute_r_rs_b_lmm` function in SABER.

```{r}
# r_frac <- c("r_rs_b_saccharina" = 0.54, "r_rs_b_cca" = 0.136, "r_rs_b_gravel" = 0.015)
r_frac <- c("r_rs_b_saccharina" = 1)#0.265, "r_rs_b_cca" = 0.011, "r_rs_b_gravel" = 0.029)

r_rs_b_saccharina <- compute_r_rs_b_lmm(r_frac)

plot_ly(x = wavelength) %>%
  add_lines(y = r_rs_b_saccharina)
```

## Inversion

The inversion process consist in finding the optimal parameter vector that can be feed to the forward model, giving approximately exact $R_\mathrm{rs}$ used as input. Here, we have used two different approaches for the inversion: 

1. A deterministic approach based on gradient-based/Newtonian optimization (L-BFGS-B/CG, etc... method) to minimize an objective function (log-likelihood default here).

2.a Bayesian framework to perform the inversion, where the posterior distribution of the parameters is sampled using  Markov-Chain Monte-Carlo (MCMC) sampling, informed with a proper prior and model likelihood.

### Deterministic inversion

The following chunk shows an illustration of a SABER inversion using gradient-based optimization method (L-BFGS-B) to minimize a log-likelihood objective function. The parameters to be inverted are passed in `par_to_inverse`, while the known parameters are fixed to known values in `par_fixed`. The boundaries of the prior/posterior are defined in `lower_b` and `upper_b`. The initial values for the parameters to be inverted are passed in `init_val`. 

```{r}
par_fixed <- c(
  "water_type" = 2,
  "theta_view" = 0,
  "theta_sun" = 20,
  #"a_g_s" = 0.017,
  "h_w" = NULL,
  "r_rs_b_saccharina" = NULL,
  "r_rs_b_cca" = NULL,
  "r_rs_b_gravel" = NULL
)


par_to_inverse <- c(#"theta_sun",
  "chl", "a_g_440", "a_g_s",
  #"a_g_s_d",
  "bb_p_550", "bb_p_gamma",

 # "a_nap_440",  "a_nap_sd",

  # "h_w",  "r_rs_b_saccharina" , "r_rs_b_cca" , "r_rs_b_gravel" ,

  "sd"
)

lower = c(0.5, 0.1, 0.005, #0.0002,
          0.002, 0.2,
          # 0.5, 0.1, 0.1, 0.1,
          0.0001
)

best = c(3, 0.75, 0.017, #0.0015,
        0.005, 0.46,
         #  4, 0.5, 0.1, 0.25,
         0.1
)

upper = c(30, 1.5, 0.022, #0.0025,
          0.015, 1,
          #  12, 1, 1, 1,
          10
)

test_inverse_grad <- inverse_gradient(
  rrs = rrs_vec,
  forward_model = "am03",
  objective_fct = "log-ll",
  optim_mtd = "L-BFGS-B",
  par_inversed = par_to_inverse,
  par_fixed = par_fixed,
  lower_b  = lower,
  init_val = best,
  upper_b  = upper,
  verbose = TRUE
)
print(test_inverse_grad)
```

### MCMC inversion

The following chunk shows an illustration of a SABER inversion using uniform prior (pass `NULL` in prior) and Gaussian likelihood. user-defined priors and random samplers can be obtained with `make_prior` and `make_prior_bundle` functions and then can be passed to the `inverse_mcmc` function. For rest of the arguements, the parameters to be inverted are passed in `par_to_inverse`, while the known parameters are fixed to known values in `par_fixed`. The boundaries of the prior/posterior are defined in `lower` and `upper`. The MCMC is run for 30,000 iterations with a burn-in of 5,000. The Differential Evolution sampler with snooker update (`"DEzs"`) is used here, but other samplers from the `BayesianTools` package can be used as well. 

```{r}
test_inverse_mcmc = inverse_mcmc(
  rrs = rrs_vec,
  forward_model = "am03",
  par_inversed = par_to_inverse,
  prior = NULL,
  lower = lower,
  best = NULL,
  upper = upper,
  par_fixed = par_fixed,
  iterations = 30000,
  burnin = 5000,
  sampler = "DEzs"
)
print(test_inverse_mcmc)
```
### Checking inversion with AOP reconstruction using inversed parameters

The SABER inversion estimated parameters are plugged into the `iop_from_oac` function and then into `forward_am03` function to simulate the $R_\mathrm{rs}$ and compare with the input one.

```{r}
#AOP reconstruction with MCMC inversed parameters
oac_mcmc <-test_inverse_mcmc[grep("sd$", names(test_inverse_mcmc), invert = T)]

iop_mcmc <- iop_from_oac(
  rrs_vec$wavelength,
  oac_mcmc
)

rrs_mcmc <- forward_am03(
  wavelength = wavelength,
  iop = iop_mcmc,
  water_type = test_case_vec["water_type"],
  theta_view = test_case_vec["theta_view"],
  theta_sun = test_case_vec["theta_sun"],
  
  h_w = NULL,
  r_b = NULL
)

#AOP reconstruction with deterministic inversed parameters
oac_grad <-test_inverse_grad$estimate[grep("sd$", (test_inverse_grad$name), invert = T)]

iop_grad <- iop_from_oac(
  rrs_vec$wavelength,
  oac_grad
)

rrs_grad <- forward_am03(
  wavelength = wavelength,
  iop = iop_grad,
  water_type = test_case_vec["water_type"],
  theta_view = test_case_vec["theta_view"],
  theta_sun = test_case_vec["theta_sun"],
  
  h_w = NULL,
  r_b = NULL
)

ply <- plot_ly() %>%
  add_lines(x = rrs_vec$wavelength, y = rrs_vec$rrs_0m, name = "rrs_input") %>%
  add_lines(x = rrs_vec$wavelength, y = rrs_mcmc, name = "rrs_retrieved_mcmc") %>%
  add_lines(x = rrs_vec$wavelength, y = rrs_grad, name = "rrs_retrieved_grad")

ply
```

